A study on student performance, engagement and experience with Kaggle InClass data challenges

### 2021-01-07 paper-kaggle.Rmd
done final proofing

---
title: A study on student performance, engagement and experience with Kaggle-in-class data challenges

@@

```{r crossperformances, echo=FALSE, fig.cap="\\label{fig:crossperformances} Students performance in classification and regression questions by competition type.", fig.width=6,fig.height=5  }

@@

ggplot() +
      annotate("rect", xmin = 0, xmax = 0.99, ymin = 1.01, ymax = 2, fill = "yellow", alpha = 0.25) +
      annotate("rect", xmin = 1.01, xmax = 2, ymin = 0, ymax = 0.99, fill = "yellow", alpha = 0.25) +
      geom_point(data=UniMelbData, aes(x=perf_class, y=perf_reg, color = Proj_alloc), size=3, alpha=0.8) +
      ylab("Regression questions performance") + 
      xlab("Classification questions performance") +
      xlim(0,2) + ylim(0, 2) +
      scale_color_brewer(palette="Dark2", name="Competition", 
                          breaks=c("regression", "classification"),   
                          labels=c("Regression", "Classification")) +
      geom_hline(yintercept = 1, color = "gray", size=0.7) +  #linetype="dotted",
      geom_vline(xintercept = 1, color = "gray", size=0.7) +
      annotate("text", x = c(0.3,0.32,1.7,1.61), y=c(0.1,1.85,1.95,0.18), 
               label = c(  a,  #(1,1)
                           b,  # "(2,2)"
                           c,  #"(3,3)"
                           d )) +
  theme_bw() +
        theme(legend.position="bottom") 
#"(4,4)"
      # annotate("text", x=0.1, y=1.8, label= "(,)")

@@

Figure \ref{fig:crossperformances} presents students' scores for classification and regression questions. A score over 1 is considered as outperform (relative to the expectation). Quarters one and three include students that underperform or outperform on both types of questions, respectively. In both cases the number of students that participated in the classification competition is very close to the number of students that participated in the regression competition (excluding a few regression students on the border of score 1). Students in quarters two and four outperform on one type of questions but not on the other type. We can see that more regression students outperform on regression questions than classification students 
(`r filter(t, Proj_alloc == "regression" ,     outperf_class == FALSE,  outperf_reg == TRUE)["n"]`<!-- 12 --> vs. 
 `r filter(t, Proj_alloc == "classification" , outperf_class == FALSE,  outperf_reg == TRUE)["n"]`<!-- 7 -->). 
Similarly, classification students do better on classification questions 
(`r filter(t, Proj_alloc == "classification" , outperf_class == TRUE,  outperf_reg == FALSE)["n"]`<!-- 11 --> vs. 
 `r filter(t, Proj_alloc == "regression" ,     outperf_class == TRUE,  outperf_reg == FALSE)["n"]`<!-- 3 -->). 
This is another evidence towards positive influence of the data competition on student's performances.

@@

Figure \ref{fig:likert} shows the survey responses related to the kaggle competition, for CSDM and ST-PG. The response rate for CSDM was 55%, with 34 of 61 students completing the survey. 
The response rate for ST-PG was 50%, 17 students out of 34 completed the survey. Overwhelmingly, students reported that they found the competition interesting and helpful for their learning in the course.

Students mostly agree that taking part in data competition improved their understanding of the covered material (Q3) and their skills to apply the covered material to real problems (Q5). Surprisingly, they didn't perceive the Kaggle challenge to be related to exam performance (Q4). Perhaps the link between the two should be emphasised by the teaching stuff when the competition is presented to students. We believe that by increasing students' awareness of the linkage between the knowledge obtained from the data competition, better understanding of the material and better marks, further increases students' engagement with the competition. This, in turn, will positively reflect on their exam scores.

>>>>

---
title: A study on student performance, engagement and experience with Kaggle InClass data challenges

@@

```{r crossperformances, echo=FALSE, fig.cap="\\label{fig:crossperformances} Student performance in classification and regression questions by competition type. The most interesting information is in the top left and bottom right quarters, where student outperform on one type of questions but not on the other type. There are more regression competition students who outperform on regression, and conversely for the classification competition students.", fig.width=6,fig.height=5  }

@@

ggplot() +
      annotate("rect", xmin = 0, xmax = 0.99, ymin = 1.01, ymax = 2, fill = "yellow", alpha = 0.25) +
      annotate("rect", xmin = 1.01, xmax = 2, ymin = 0, ymax = 0.99, fill = "yellow", alpha = 0.25) +
      geom_hline(yintercept = 1, color = "gray", size=0.7) +  #linetype="dotted",
      geom_vline(xintercept = 1, color = "gray", size=0.7) +
      geom_point(data=UniMelbData, aes(x=perf_class, y=perf_reg, color = Proj_alloc), size=3, alpha=0.8) +
      ylab("Regression questions performance") + 
      xlab("Classification questions performance") +
      xlim(0,2) + ylim(0, 2) +
      scale_color_brewer(palette="Dark2", name="Competition", 
                          breaks=c("regression", "classification"),   
                          labels=c("Regression", "Classification")) +
      annotate("text", x = c(0.3,0.32,1.7,1.61), y=c(0.1,1.85,1.95,0.18), 
               label = c(  a,  #(1,1)
                           b,  # "(2,2)"
                           c,  #"(3,3)"
                           d )) +
  theme_bw() +
        theme(legend.position="bottom") 
#"(4,4)"
      # annotate("text", x=0.1, y=1.8, label= "(,)")

@@

Figure \ref{fig:crossperformances} presents student scores for classification and regression questions. A score over 1 is considered as outperform (relative to the expectation). Quarters one and three include students that underperform or outperform on both types of questions, respectively. In both cases the number of students that participated in the classification competition is very close to the number of students that participated in the regression competition (excluding a few regression students on the border of score 1). Students in top left and bottom right quarters outperform on one type of questions but not on the other type. We can see that more regression students outperform on regression questions than classification students 
(`r filter(t, Proj_alloc == "regression" ,     outperf_class == FALSE,  outperf_reg == TRUE)["n"]`<!-- 12 --> vs. 
 `r filter(t, Proj_alloc == "classification" , outperf_class == FALSE,  outperf_reg == TRUE)["n"]`<!-- 7 -->). 
Similarly, classification students do better on classification questions 
(`r filter(t, Proj_alloc == "classification" , outperf_class == TRUE,  outperf_reg == FALSE)["n"]`<!-- 11 --> vs. 
 `r filter(t, Proj_alloc == "regression" ,     outperf_class == TRUE,  outperf_reg == FALSE)["n"]`<!-- 3 -->). 
This is another evidence towards positive influence of the data competition on student's performances.

@@

Figure \ref{fig:likert} shows the survey responses related to the Kaggle competition, for CSDM and ST-PG. The response rate for CSDM was 55%, with 34 of 61 students completing the survey. 
The response rate for ST-PG was 50%, 17 students out of 34 completed the survey. Overwhelmingly, students reported that they found the competition interesting and helpful for their learning in the course.

Students mostly agree that taking part in the data competition improved their learning experience, especially understanding of the covered material (Q3) and their skills to apply the covered material to real problems (Q5). Surprisingly, less students perceived the Kaggle challenge might help with exam performance (Q4). Perhaps the link between the two could be emphasised by instructors when the competition is presented to students. Increasing student awareness of the association between the knowledge obtained from the data competition, better understanding of the material and better marks, might increase all students engagement with the competition.
